---
title: "tiic_analysis"
output: html_document
date: "2024-07-10"
---

```{r setup, include=FALSE, warnings=FALSE}
knitr::opts_chunk$set(echo = FALSE, warnings=FALSE)
library(tidyverse)
library(survival)
library(broom)
library(janitor)
library(meta)
library(mice)
library(lemon)
library(patchwork)
library(tableone)
library(scales)
```

## Read in case data and TIIC automated scores 

```{r data}
# Phenotype data
pheno <- read_csv("./Data/case_data.csv")
# TIIC score data by core
halo <- read_csv("./Data/tiic_scores.csv") 

# Supplementary table 3
cores_case <- halo |> select(study, marker, bcac_id) |> 
  group_by(marker, bcac_id) |> summarise(count = n()) |> ungroup() |> 
  count(marker, count) |>   pivot_wider(names_from = 'marker', values_from='n')

write_csv(cores_case, "./Outputs/table_s3.csv")

# Supplementary table 4
cores_study <- halo |> select(study, marker, bcac_id) |> 
  group_by(marker, study) |> summarise(count = n()) |> ungroup() |> 
  pivot_wider(names_from = 'marker', values_from='count')
write_csv(cores_study, "./Outputs/table_s4.csv")

# Supplementary table 8
mortality <- pheno |> filter(yod>=2000) |> select(status, fu_15, br_15, bcss_15) |>
  mutate(fu_gp = cut(fu_15, breaks=c(0, 5, 10, 15), right=FALSE))

bcss <- count(filter(mortality, bcss_15==1), fu_gp, bcss_15) |> select(1, 3) |> rename(bcss = n)
br <- count(filter(mortality, br_15==1), fu_gp, br_15) |> select(1, 3) |> rename(br = n)
all <- count(filter(mortality, status==1), fu_gp, status) |> select(1, 3) |> rename(all = n) |>
  left_join(bcss) |> left_join(br) |>
  mutate(other = all - br,
         known = bcss + all - br,
         unknown = br - bcss) |>
  filter(!is.na(fu_gp)) |> select(fu_gp, bcss, other, known, unknown)
write_csv(all, "./Outputs/table_s8.csv")
rm(cores_case, cores_study, mortality, bcss, br, all)

```

## Functions for data wrangling and modelling

data_fun generates patient level scores from the automated core level scores as a function of the threshold for excluding cores based on minimum tissue area.  The scores are then merged with the clinical data and the derived age variables created.

```{r functions}
# Function to generate patient level data for different thresholds of viable core area
data_fun <- function(x) {
  data <- halo |>
    filter(area_all_mn>x) |>
    #Generate patient level scores
    group_by(marker, bcac_id) |> 
    summarise(cores=n(),
              area_all = mean(area_all_mn, na.rm=TRUE),
              area_stroma = mean(area_stroma_mn, na.rm=TRUE),
              area_tumour = mean(area_tumour_mn, na.rm=TRUE),
              area_artefact = mean(area_artefact_mn, na.rm=TRUE),
              area_til_all = mean(area_all_til_halo, na.rm=TRUE),
              area_til_stroma = mean(area_stroma_til_halo, na.rm=TRUE),
              area_til_tumour = mean(area_tumour_til_halo, na.rm=TRUE)) |>
    # Generate patient level percent area scores
    mutate(pct_til_all = area_til_all/(10^4*area_all),
           pct_til_stroma = ifelse(area_stroma==0, NA, area_til_stroma/(area_stroma*10^4)),
           pct_til_tumour = ifelse(area_tumour==0, NA, area_til_tumour/(area_tumour*10^4)),
           pct_artefact = area_artefact/(area_artefact + area_all),
           log_pct_til_all = log(pct_til_all + 0.01),
           log_pct_til_stroma = log(pct_til_stroma + 0.01), 
           log_pct_til_tumour = log(pct_til_tumour + 0.01), 
           std_all = scale(log(pct_til_all + 0.1), center=TRUE, scale=TRUE)[,1],
           std_stroma = scale(log(pct_til_stroma + 0.1), center=TRUE, scale=TRUE)[,1],
           std_tumour = scale(log(pct_til_tumour + 0.1), center=TRUE, scale=TRUE)[,1]) |>
    select(bcac_id, marker, cores, 11:13, 18:20) |>
    pivot_wider(id_cols = bcac_id,
                names_from=marker,
                values_from = c(cores, pct_til_all, pct_til_stroma, pct_til_tumour,
                                std_all, std_stroma, std_tumour)) |>
    right_join(filter(pheno, !is.na(er), !is.na(br_15))) |>
    mutate(threshold=x,
           # Generate PREDICT age variables
           age1 = ifelse(er==1, (age_diag/10)^-2, age_diag),
           age2 = ifelse(er==1, (age_diag/10)^-2*log(age_diag/10), 0)) |>
    select(-er_ast, -er_ari)
  return(data)
}

models_fun <- function(models) {
  summaries <- lapply(models, summary)
  term <- names(df[,6:29])
  estimate <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,1])))
  std.error <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,3])))
  p.value <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,5])))
  n <- as.numeric(unlist(lapply(summaries, function(x) x$n)))
  nevents <- as.numeric(unlist(lapply(summaries, function(x) x$nevent)))
  uv <- tibble(term, estimate, std.error, p.value, n, nevents) |>
    mutate(imputation = "Not applicable",
           variable = "Single marker",
           area = threshold,
           subtype = ifelse(er_status==1, "ER positive", "ER negative"),
           analysis = "Complete case analysis")
  return(uv)
}
```

## Single marker complete case analysis to compare percent area with log percent and to compare different area exclusion thresholds

```{r cca}
# Vector of thresholds for exluding cores based on minimum tissue area
t_values <- c(0.05, 0.1, 0.15, 0.2, 0.25, 0.3)

data <- map_df(t_values, data_fun) |>
  select(yod, threshold, er, site, enter, fu_15, br_15, 
         age1, age2, size, nodes, grade, pct_til_all_CD163:std_tumour_FOXP3) |> 
  pivot_longer(cols=13:36, names_to='tiic', values_to='score')

# Set up empty tibble for storing results
uni_marker <- tibble()

df <- data |> group_by(threshold, er, tiic) |> nest()

partial <- df |> 
  mutate(model = map(data, ~coxph(Surv(enter, fu_15, br_15) ~ site + age1 + age2 + score, data=.x))) |>
  mutate(tidy.model = map(model, broom::tidy),
         number = map_dbl(model, ~(.x$n)),
         events = map_dbl(model, ~(.x$nevent))) |>
  ungroup() |>
  select(threshold, er, tiic, number, events, tidy.model) |>
  unnest(cols=c(tidy.model)) |>
  filter(term == "score") |>
  mutate(adjust="Partial")

uni_marker <- df |> 
  mutate(model = map(data, ~coxph(Surv(enter, fu_15, br_15) ~ site + age1 + age2 + 
                                    size + nodes + grade + tt(grade) + score, 
                                  tt = function(x, t, ...) { x * log(t)}, data=.x))) |>
  mutate(tidy.model = map(model, broom::tidy),
         number = map_dbl(model, ~(.x$n)),
         events = map_dbl(model, ~(.x$nevent))) |>
  ungroup() |>
  select(threshold, er, tiic, number, events, tidy.model) |>
  unnest(cols=c(tidy.model)) |>
  filter(term == "score") |>
  mutate(adjust="Full") |>
  bind_rows(partial) |>
  mutate(score = ifelse(str_detect(tiic, "pct"), "Pct", "Std"),
         marker = case_when(str_detect(tiic, "CD8") ~ "CD8",
                           str_detect(tiic, "CD20") ~ "CD20",
                           str_detect(tiic, "CD163") ~ "CD163",
                           str_detect(tiic, "FOXP3") ~ "FOXP3"),
         compartment = case_when(str_detect(tiic, "all") ~ "All",
                                 str_detect(tiic, "stroma") ~ "Stroma",
                                 str_detect(tiic, "tumour") ~ "Tumour"),
         hr = exp(estimate),
         lcl = exp(estimate - qnorm(.975)*std.error),
         ucl = exp(estimate + qnorm(.975)*std.error),
         signif = p.value<0.001,
         variable="Single marker",
         analysis="Complete case analysis",
         subtype = ifelse(er==0, "ER negative", "ER positive")) |>
  rename(area = threshold) |> select(variable, analysis, area, subtype, marker, compartment, adjust, score, 
         estimate, std.error, hr, lcl, ucl, p.value, signif, number, events) |>
  arrange(variable, analysis, area, subtype, marker, compartment, adjust, score)

write_csv(uni_marker, "./Outputs/unimarker_results.csv")
rm(partial, df, data)
```

## Choosing best threshold for year of diagnosis, tumour area threshold and TIIC score transformation

```{r}
# Count which score (percent area of log percent) is best for each model
uni_marker |> group_by(area, subtype, marker, compartment, adjust) |>
  slice(which.min(p.value)) |> ungroup() |> count(score)

# Count which area threshold is best for each model
uni_marker |> group_by(subtype, marker, compartment, adjust, score) |>
  slice(which.min(p.value)) |> ungroup() |> count(area)
```

Area threshold of greater than 0.05 mm2 is the most often the most significant association for any given analysis, and the standardised log percentage is most often the most significant association for any given analysis

# Imputation of missing data  ----

Missingness of TIIC scores depends on threshold used for excluding cores with limited tissue area.  Thus data needs to be imputed for each threshold for data exclusion.

```{r impute}
# Number of imputed data sets
iterations <- 10
df <- tibble()

for (threshold in t_values) { 
  mv <- halo |>
    mutate(area_all_mn = ifelse(area_all_mn<threshold, NA, area_all_mn),
           area_stroma_mn = ifelse(area_all_mn<threshold, NA, area_stroma_mn),
           area_tumour_mn = ifelse(area_all_mn<threshold, NA, area_tumour_mn),
           area_all_til_halo = ifelse(area_all_mn<threshold, NA, area_all_til_halo),
           area_stroma_til_halo = ifelse(area_all_mn<threshold, NA, area_stroma_til_halo),
           area_tumour_til_halo = ifelse(area_all_mn<threshold, NA, area_tumour_til_halo)) |>
    #Generate patient level scores
    group_by(marker, bcac_id) |>
    summarise(cores=n(),
              area_all = mean(area_all_mn, na.rm=TRUE),
              area_stroma = mean(area_stroma_mn, na.rm=TRUE),
              area_tumour = mean(area_tumour_mn, na.rm=TRUE),
              area_artefact = mean(area_artefact_mn, na.rm=TRUE),
              area_til_all = mean(area_all_til_halo, na.rm=TRUE),
              area_til_stroma = mean(area_stroma_til_halo, na.rm=TRUE),
              area_til_tumour = mean(area_tumour_til_halo, na.rm=TRUE)) |>
    mutate(pct_til_all = area_til_all/(10^4*area_all),
           pct_til_stroma = ifelse(area_stroma==0, NA, area_til_stroma/(area_stroma*10^4)),
           pct_til_tumour = ifelse(area_tumour==0, NA, area_til_tumour/(area_tumour*10^4)),
           pct_artefact = area_artefact/(area_artefact + area_all),
           log_pct_til_all = log(pct_til_all + 0.1),
           log_pct_til_stroma = log(pct_til_stroma + 0.1),
           log_pct_til_tumour = log(pct_til_tumour + 0.1)) |>
    select(bcac_id, marker, cores, 15:17 ) |>
    pivot_wider(id_cols = bcac_id,
                names_from=marker,
                values_from = c(cores, log_pct_til_all, log_pct_til_stroma, log_pct_til_tumour)) |>
    right_join(pheno) |> 
    # exclude variables not used for imputation
    select(-site, -bcss_15) |> 
    mutate(nodes = log(nodes+0.1),
           size = log(size),
           grade= as.factor(grade),
           er = as.factor(er),
           pr = as.factor(pr),
           her2 = as.factor(her2),
           horm = as.factor(horm),
           tras=as.factor(tras),
           radio=as.factor(radio),
           chemo=as.factor(chemo))
  
  set.seed <- 290962
  imp <- mice(mv, m=iterations, defaultMethod = c("norm", "logreg", "polyreg", "polr"))
  imp.long <- mice::complete(imp,"long",include = T)  |>
    mutate(grade = as.numeric(grade),
           size = exp(size),
           nodes = case_when(exp(nodes)-0.1 < 1e-05 ~ 0,
                             exp(nodes)-0.1 > 50 ~ 50,
                             TRUE ~ exp(nodes)-0.1),
           age1 = ifelse(er==1, (age_diag/10)^-2, age_diag),
           age2 = ifelse(er==1, (age_diag/10)^-2*log(age_diag/10), 0),
           threshold = threshold) |>
    group_by(.imp) |>
    mutate(  all_CD8 = scale(log_pct_til_all_CD8, center=TRUE, scale=TRUE)[,1],
             stroma_CD8 = scale(log_pct_til_stroma_CD8, center=TRUE, scale=TRUE)[,1],
             tumour_CD8 = scale(log_pct_til_tumour_CD8, center=TRUE, scale=TRUE)[,1],
             all_CD20 = scale(log_pct_til_all_CD20, center=TRUE, scale=TRUE)[,1],
             stroma_CD20 = scale(log_pct_til_stroma_CD20, center=TRUE, scale=TRUE)[,1],
             tumour_CD20 = scale(log_pct_til_tumour_CD20, center=TRUE, scale=TRUE)[,1],
             all_CD163 = scale(log_pct_til_all_CD163, center=TRUE, scale=TRUE)[,1],
             stroma_CD163 = scale(log_pct_til_stroma_CD163, center=TRUE, scale=TRUE)[,1],
             tumour_CD163 = scale(log_pct_til_tumour_CD163, center=TRUE, scale=TRUE)[,1],
             all_FOXP3 = scale(log_pct_til_all_FOXP3, center=TRUE, scale=TRUE)[,1],
             stroma_FOXP3 = scale(log_pct_til_stroma_FOXP3, center=TRUE, scale=TRUE)[,1],
             tumour_FOXP3 = scale(log_pct_til_tumour_FOXP3, center=TRUE, scale=TRUE)[,1]) |>
    ungroup() |>
    select(.imp:cores_FOXP3, study:tumour_FOXP3)
  
  df <- bind_rows(df, imp.long)
  
}

df <- df |> 
  select(.imp, .id, study, bcac_id, threshold, enter, fu_15, br_15, age1, age2, grade:her2,
         chemo, all_CD8:tumour_FOXP3 )
sites <- tibble()

for (cutoff in t_values) {
  for (iteration in 0:10) {
  events <- df |>
    filter(.imp==iteration, threshold==cutoff) |>  
    group_by(study, er) |> summarise(events=sum(br_15, na.rm=TRUE))

  er0_gp <- as.vector(events |> filter(er==0, events<10) |> select(1))$study
  er1_gp <- as.vector(events |> filter(er==1, events<10) |> select(1))$study
  diam6 <- c('ABC', 'MBC', 'MMH', 'NEA', 'OFB', 'ORI', 'RBC', 'SEA', 'SBC', 'SKK', 'SZB' )
  diam1 <- c('EPI', 'SGB','BGS', 'CPS', 'EST', 'PSH', 'GES', 'MAR', 'PBC', 'PLC', 'UNC')
  df2 <- df |> filter(.imp==iteration, threshold==cutoff) |>
    select(.imp, threshold, study, bcac_id, er) |>
    mutate(site0 = case_when(study %in% er0_gp & study %in% diam6 ~ "DIA6",
                                        study %in% er0_gp & study %in% diam1 ~ "DIA1",
                                        TRUE ~ study),
           site1 = case_when(study %in% er1_gp & study %in% diam6 ~ "DIA6",
                                        study %in% er1_gp & study %in% diam1 ~ "DIA1",
                                        TRUE ~ study),
           site = ifelse(er==0, site0, site1)) |> 
    select(.imp, threshold, bcac_id, site)
  sites <- sites |> bind_rows(df2)
  rm(df2)
  }
}

imp.long <- df |> left_join(sites) |> select(.imp, .id, threshold, study, site, everything())

write_csv(imp.long, "./Outputs/imputed_data.csv")
rm(df, mv, sites, events )

```

## Multi-variable Cox regression analyses  ----

After imputation the single marker and multi-marker analyses are run with partial and full adjustment for other prognostic variables.

```{r cox_mv}
# Read in imputed data and reshape 
imp.long <- read_csv("./Outputs/imputed_data.csv")

df <- imp.long |> filter(.imp!=0) |>
  pivot_longer(all_CD8:tumour_FOXP3, names_to='tiic', values_to='score') |>
  # Nest the data by imputation, area thrshold, ER status and TIIC marker
  group_by(.imp, threshold, er, tiic) |> nest()

# Partially adjusted single marker models (10*6*2*12=1440 models)
results_i <- df |> 
  mutate(model = map(data, ~coxph(Surv(enter, fu_15, br_15) ~ site + age1 + age2 + score, data=.x))) |>
  mutate(tidy.model = map(model, broom::tidy),
         number = map_dbl(model, ~(.x$n)),
         events = map_dbl(model, ~(.x$nevent))) |>
  ungroup() |>
  select(.imp, threshold, er, tiic, number, events, tidy.model) |>
  unnest(cols=c(tidy.model)) |>
  filter(term == "score") |>
  mutate(analysis = "Multiple imputation",
         variable="Single marker",
         adjust="Partial")

# Run fully adjusted model
# Unnest data and re-nest without imputation (memory management for fully adjusted model)
df <- imp.long |> filter(.imp!=0) |>
  pivot_longer(all_CD8:tumour_FOXP3, names_to='tiic', values_to='score') |>
  # Nest the data by imputation, area thrshold, ER status and TIIC marker
  group_by(threshold, er, tiic) |> nest()

# Use loop to loop through each imputation
for (iteration in 1:10) {
  full <- df |>
    mutate(model = map(data, ~coxph(Surv(enter, fu_15, br_15) ~ site + age1 + age2 +
                                      size + nodes + grade + tt(grade) + score,
                                    tt = function(x, t, ...) { x * log(t)}, data=filter(.x, .imp==iteration)))) |>
    mutate(tidy.model = map(model, broom::tidy),
           number = map_dbl(model, ~(.x$n)),
           events = map_dbl(model, ~(.x$nevent))) |>
    ungroup() |>
    select(threshold, er, tiic, number, events, tidy.model) |>
    unnest(cols=c(tidy.model)) |>
    filter(term == "score") |>
    mutate(.imp=iteration,
           analysis = "Multiple imputation",
           variable="Single marker",
           adjust="Full") #|>
    results_i <- bind_rows(results_i, full)
    rm(full)
    print(iteration)
}

results_i <- results_i |>
  mutate(marker= case_when(str_detect(tiic, "CD8") ~ "CD8",
                    str_detect(tiic, "CD20") ~ "CD20",
                    str_detect(tiic, "CD163") ~ "CD163",
                    str_detect(tiic, "FOXP3") ~ "FOXP3"),
         compartment = case_when(str_detect(tiic, "all") ~ "all",
                        str_detect(tiic, "stroma") ~ "stroma",
                        str_detect(tiic, "tumour") ~ "tumour")) |>
  select(-tiic, -term)

# Run multi-marker models
df <- imp.long |> filter(.imp!=0) |>
  pivot_longer(all_CD8:tumour_FOXP3, names_to = c("compartment", ".value"), names_sep="_" ) |> 
  # Nest data
  group_by(.imp, threshold, er, compartment) |> nest()

multi.partial <- df |> 
  mutate(model = map(data, ~coxph(Surv(enter, fu_15, br_15) ~ site + age1 + age2 + 
                                     CD8 + CD20 + CD163 + FOXP3, 
                                   data=.x))) |>
  mutate(tidy.model = map(model, broom::tidy),
         number = map_dbl(model, ~(.x$n)),
         events = map_dbl(model, ~(.x$nevent))) |>
  ungroup() |>  
  select(.imp, threshold, er, compartment, number, events, tidy.model) |>
  unnest(cols=c(tidy.model)) |> 
  filter(term %in% c("CD8", "CD20", "CD163", "FOXP3")) |>
  mutate(analysis = "Multiple imputation",
         variable="Multi-marker",
         adjust="Partial") |>
  rename(marker=term)

results_i <- bind_rows(results_i, multi.partial)
rm(multi.partial)

df <- imp.long |> filter(.imp!=0) |>
  pivot_longer(all_CD8:tumour_FOXP3, names_to = c("compartment", ".value"), names_sep="_" ) |> 
  # Nest data
  group_by(threshold, er, compartment) |> nest()

for (iteration in 1:10) {
  multi.full <- df |> 
    mutate(model = map(data, ~coxph(Surv(enter, fu_15, br_15) ~ site + age1 + age2 + 
                                      size + nodes + grade + tt(grade) + CD8 + CD20 + CD163 + FOXP3, 
                                    tt = function(x, t, ...) { x * log(t)}, data=filter(.x, .imp==iteration)))) |>
    mutate(tidy.model = map(model, broom::tidy),
           number = map_dbl(model, ~(.x$n)),
           events = map_dbl(model, ~(.x$nevent))) |>
    ungroup() |>
    select(threshold, er, compartment, number, events, tidy.model) |>
    unnest(cols=c(tidy.model)) |>
    filter(term %in% c("CD8", "CD20", "CD163", "FOXP3")) |>
    mutate(.imp=iteration, 
           analysis = "Multiple imputation",
           variable="Multi-marker",
           adjust="Full") |>
    rename(marker=term)

  results_i <- bind_rows(results_i, multi.full)
  rm(multi.full)
  print(iteration)
}

rm(df)

```

## Combine the results across analysis of each imputed dataset 

Results from analysis of 10 imputed data sets for each exclusion threshold are combined using Rubin's rules and then output saved.

```{r rubin}
est_mean <- results_i |>
  group_by(variable, er, marker, compartment, threshold, adjust) |>
  summarise(est_mean=mean(estimate))

results  <-  results_i |>
  left_join(est_mean) |>
  mutate(diff = estimate - est_mean,
         variance=std.error^2) |>
  group_by(analysis, variable, er, marker, compartment, threshold, adjust) |>
  rename(area = threshold) |>
  summarise(estimate=mean(estimate),
            var_w=mean(variance),
            var_b = mean(diff^2)/9,
            number = round(mean(number)),
            events=round(mean(events))) |>
  mutate(score = "Std",
         std.error = (var_w + var_b + var_b/10)^.5,
         statistic = estimate/std.error,
         p.value = 2*pnorm(abs(statistic), lower.tail=F),
         signif = p.value<0.05,
         hr = exp(estimate),
         lcl = exp(estimate - qnorm(0.975)*std.error),
         ucl = exp(estimate + qnorm(0.975)*std.error),
         subtype = ifelse(er==0, "ER negative", "ER positive")) |>
  select(-var_w, -var_b, -statistic) |>
  arrange(er, marker, compartment, desc(variable), desc(adjust)) |> ungroup()

results <- results |> bind_rows(uni_marker) |>
  select(variable, analysis, area, subtype, marker, compartment, adjust, score, estimate, std.error:ucl, number, events) |> arrange(desc(variable), analysis, area, subtype, marker, compartment, adjust)

rm(est_mean, results_i, uni_marker)

write_csv(results, "./Outputs/table_s5.csv")

```

## Fraction model variance explained by TIIC scores

The relative variation [21] accounted for by TIIC scores as 
1 – adequacy index = 1 – LRstandard/LRstandard+B
where LR is the likelihood ratio chi-squared statistic, the standard model includes age at diagnosis, ER status, grade, tumor size, number of positive nodes and B denotes the inclusion of TIIC scores as predictors.

Mean value across 10 data sets imputed with the tissue adequacy threshold set at .05mm2. 

```{r variance, echo=FALSE}
#imp.long <- read_csv("./Outputs/imputed_data.csv") |> filter(threshold==0.05)

adequacy <- tibble(imputation= numeric(),
                   er = numeric(),
                   tils= numeric(),
                   site= numeric(),
                   age = numeric(),
                   size = numeric(),
                   nodes = numeric(),
                   grade = numeric(),
                   all = numeric())

ll_fun <- function(x, y) {
  df <- filter(imp.long, er==x, .imp==y) 
  ll <- c(0, 0, 0, 0, 0, 0 ,0)
  i <- 0
  for (variables in c("site + age1 + age2 + size + nodes + grade + tt(grade)",
                      "age1 + age2 + size + nodes + grade + tt(grade) + all_CD163 + all_CD20 + 
                       all_CD8 + all_FOXP3",
                       "site + size + nodes + grade + tt(grade) + all_CD163 + all_CD20 + 
                       all_CD8 + all_FOXP3",
                      "site + age1 + age2 + nodes + grade + tt(grade) + all_CD163 + all_CD20 + 
                       all_CD8 + all_FOXP3",
                       "site + age1 + age2 + size + grade + tt(grade) + all_CD163 + all_CD20 + 
                       all_CD8 + all_FOXP3",
                      "site + age1 + age2 + size + nodes + all_CD163 + all_CD20 + 
                       all_CD8 + all_FOXP3",
                      "site + age1 + age2 + size + nodes + grade + tt(grade) + all_CD163 + all_CD20 + 
                       all_CD8 + all_FOXP3")) {
    formula <- paste0("Surv(enter, fu_15, br_15) ~ ", variables) 
    fit <- coxph(as.formula(formula),
                 data=df, 
                 tt = function(x, t, ...) { x * log(t)})
    temp <- 2*(fit$loglik[2]-fit$loglik[1])
    i <- i + 1
    ll[i] <- temp
  }
  return(ll)
}

for (imputation in 1:10) {
  for (er in 0:1) {
    ll = ll_fun(x=er, y=imputation)
    row <- tibble(imputation= imputation, 
                  er = er, 
                  tils= ll[1],
                  site = ll[2],
                  age = ll[3],
                  size = ll[4],
                  nodes = ll[5],
                  grade = ll[6],
                  all = ll[7]
                  )
    adequacy <- bind_rows(adequacy, row)
  }
}

adequacy_summary <- adequacy |> filter(!is.na(site)) |>
  mutate(rsq_tils = 1 - tils/all,
         rsq_site = 1 - site/all,
         rsq_age = 1 - age/all,
         rsq_size = 1 - size/all,
         rsq_nodes = 1 - nodes/all,
         rsq_grade = 1 - grade/all) |>
  group_by(er) |>
  summarise(rsq_tils = mean(rsq_tils),
            rsq_site = mean(rsq_site),
            rsq_age = mean(rsq_age),
            rsq_size = mean(rsq_size),
            rsq_nodes = mean(rsq_nodes),
            rsq_grade = mean(rsq_grade))

write_csv(adequacy_summary, "./Outputs/adequacy_results.csv")
#rm(adequacy, adequacy_summary)
```

## Evaluating heterogeneity between studies

Based on the imputed data when using the 0.05 mm2 threshold

```{r}
# Generate empty object to store results
study_results_i <- tibble(study= character(),
                     imputation = numeric(),
                     er = numeric(),
                     term = character(),
                     estimate = numeric(),
                     std.error = numeric())

for (imputation in 1:10) {
  for (er_status in 0:1) {
    sites <- imp.long |> filter(er==er_status, threshold==0.05) |> select(site)
    studies <- unique(sites$site)
        for (set in studies) {
            df <- imp.long |> filter(er==er_status, .imp==imputation, threshold==0.05, site==set)
            
            fit <- coxph(Surv(enter, fu_15, br_15) ~ all_CD163 + all_CD20 + 
                       all_CD8 + all_FOXP3 + 
                           age1 + age2 + size + nodes + grade + tt(grade),
                         tt = function(x, t, ...) { x * log(t)},
                         data=df)
            
            het <- tidy(fit)[1:4, c(1:3,5)] |>
              mutate(study = set,
                     imputation = imputation,
                     er = er_status)
            study_results_i <- bind_rows(study_results_i, het)
    }
  }
}

est_mean <- study_results_i |>
  mutate(marker= case_when(str_detect(term, "CD8") ~ "CD8",
                           str_detect(term, "CD20") ~ "CD20",
                           str_detect(term, "CD163") ~ "CD163",
                           str_detect(term, "FOXP3") ~ "FOXP3"),
         compartment = "All") |>
  group_by(study, er, marker) |>
  summarise(est_mean=mean(estimate))

study_results  <-  study_results_i |>
  mutate(marker= case_when(str_detect(term, "CD8") ~ "CD8",
                           str_detect(term, "CD20") ~ "CD20",
                           str_detect(term, "CD163") ~ "CD163",
                           str_detect(term, "FOXP3") ~ "FOXP3"),
         compartment = case_when(str_detect(term, "all") ~ "All",
                                 str_detect(term, "stroma") ~ "Stroma",
                                 str_detect(term, "tumour") ~ "Tumour")) |>
  left_join(est_mean) |>
  mutate(diff = estimate - est_mean,
         variance=std.error^2) |>
  group_by(study, er, marker, compartment) |>
  summarise(estimate=mean(estimate),
            var_w=mean(variance),
            var_b = mean(diff^2)/9) |>
  mutate(std.error = (var_w + var_b + var_b/10)^.5,
         statistic = estimate/std.error,
         p.value = 2*pnorm(abs(statistic), lower.tail=F),
         analysis = "Multiple imputation",
         threshold = 0.05,
         adjust = "Full",
         variable = "Multi-marker") |>
  select(analysis, adjust, variable, threshold, marker, compartment, er, study, estimate, std.error, p.value, -var_w, -var_b) 
rm(study_results_i)

# Meta-analysis of the study specific results
meta_fun <- function(x, y) { 
  df <- study_results |> filter(er==x, marker==y)
  m <- metagen(data=df,
          studlab=study, 
          TE=estimate, seTE=std.error,
          backtransf=TRUE, sm="HR",
          comb.random=FALSE)
  return(c(m$I2, m$pval.Q))
}

het_results <- tibble(marker=character(),
                      er=numeric(),
                      p.het = numeric(),
                      i2 = numeric())

for (marker in c("CD8", "CD20", "CD163", "FOXP3")) {
  for (er in 0:1) {
    test <- meta_fun(x=er, y=marker)
    out <- tibble(marker=marker,
                  er=er,
                  p.het = test[2],
                  i2 = test[1])
    het_results <- bind_rows(het_results, out)
    rm(test, out)
  }
}
  
write_csv(het_results, "./Outputs/table_s6.csv")
rm(study_results_i, study_results, het_results)
```

## Comparing effect estimates in HER2+ v HER2 - and chemo + and chemo -

```{r subgroups, include=FALSE}
iterations <- 10
cutoff <- 0.05
#imp.long <- read_csv("./Outputs/imputed_data.csv") 

df <- imp.long |>
  filter(threshold==cutoff, .imp!=0) |>
  mutate(outcome=br_15)
  
subgroup.results <- tibble()
for (subgroup in c("her2", "chemo")) {
  df <- imp.long |>
  filter(threshold==cutoff) |>
  mutate(outcome=br_15,
         temp = subgroup,
         subset=ifelse(temp=="her2", her2, chemo))
  
  for (group in 0:1) {
    for (er_status in 0:1) {
      for (iteration in 1:iterations) {
        
        imp.subset <- df |> filter(er==er_status, .imp==iteration, subset==group) 

        models <- lapply(imp.subset[19:30],
                         function(x) coxph(Surv(enter, fu_15, outcome) ~
                                             x + age1 + age2 + nodes + size + grade + tt(grade) + site,
                                           tt = function(x, t, ...) { x * log(t)},
                                           data=imp.subset))
        summaries <- lapply(models, summary)
        term <- names(imp.subset[,19:30])
        estimate <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,1])))
        std.error <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,3])))
        p.value <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,5])))
        n <- as.numeric(unlist(lapply(summaries, function(x) x$n)))
        nevents <- as.numeric(unlist(lapply(summaries, function(x) x$nevent)))
        uv <- tibble(term, estimate, std.error, p.value, n, nevents) |>
          mutate(subset = subgroup,
                 stratum = ifelse(group==0, "G0", "G1"),
                 n = n,
                 nevents = nevents,
                 iteration=iteration,
                 variable = "Single marker",
                 threshold = cutoff,
                 subtype = ifelse(er_status==1, "ER positive", "ER negative"),
                 adjust = "Full")
        
        subgroup.results <- bind_rows(subgroup.results, uv)
        
        for (compartment in c("all", "stroma", "tumour")) { 
          
          v1 <- paste0(compartment, "_CD8 + ")
          v2 <- paste0(compartment, "_FOXP3 + ")
          v3 <- paste0(compartment, "_CD20 + ")
          v4 <- paste0(compartment, "_CD163 + ")
          
          formula <- paste0("Surv(enter, fu_15, outcome) ~ ", v1, v2, v3, v4, 
                            " + age1 + age2 + grade + size + nodes + tt(grade) +", 
                            " site")
          mi <- coxph(as.formula(formula), 
                      tt = function(x, t, ...) { x * log(t)},
                      data=imp.subset)
          mv <- tidy(mi)[1:4, c(1:3,5)] |>
            mutate(subset = subgroup,
                   stratum = ifelse(group==0, "G0", "G1"),
                   iteration = iteration,
                   variable = "Multi-marker",
                   threshold = cutoff,
                   subtype = ifelse(er_status==1, "ER positive", "ER negative"),
                   adjust = "Full")
          mv$n <- mi$n
          mv$nevents <- mi$nevent
          subgroup.results <- bind_rows(subgroup.results, mv)
        }
      }
    }
  }
} 

subgroup.results <- subgroup.results |> 
  mutate(marker= case_when(str_detect(term, "CD8") ~ "CD8",
                           str_detect(term, "CD20") ~ "CD20",
                           str_detect(term, "CD163") ~ "CD163",
                           str_detect(term, "FOXP3") ~ "FOXP3"),
         compartment = case_when(str_detect(term, "all") ~ "All",
                                 str_detect(term, "stroma") ~ "Stroma",
                                 str_detect(term, "tumour") ~ "Tumour"))
est.mean <- subgroup.results |>
  group_by(subset, stratum, variable, subtype, marker, compartment) |>
  summarise(number=n(), mean.est=mean(estimate)) |> ungroup()

subgroup.results  <-  subgroup.results |>
  left_join(est.mean) |>
  mutate(diff = estimate - mean.est,
         variance=std.error^2) |>
  group_by(subset, stratum, variable, subtype, marker, compartment) |>
  summarise(estimate=mean(estimate),
            var_w=mean(variance),
            var_b = mean(diff^2)/9,
            n = round(mean(n)),
            nevents=round(mean(nevents))) |>
  mutate(std.error = (var_w + var_b + var_b/10)^.5,
         statistic = estimate/std.error,
         p.value = 2*pnorm(abs(statistic), lower.tail=F),
         hr = exp(estimate),
         lcl = exp(estimate - qnorm(0.975)*std.error),
         ucl = exp(estimate + qnorm(0.975)*std.error)) |>
  select(-var_w, -var_b, -statistic) |> ungroup()

subs <- subgroup.results |> select(1:7, 10, 11) |>
  pivot_wider(names_from="stratum", values_from=7:9) |>
  mutate(subset = factor(subset, labels=c("Chemo v no chemo", "HER2 pos v HER2 neg")),
         diff = abs(estimate_G0-estimate_G1),
         se.diff = sqrt(std.error_G0^2 + std.error_G1^2),
         z = diff/se.diff,
         p.value.diff = 2*pnorm(z, lower.tail=FALSE)) |> arrange(p.value.diff) |>
  #filter(as.integer(subset)<3) |>
  mutate(rank = rank(p.value.diff),
         p.exp = (rank - 0.5)/96,
         significant = p.value.diff<0.05,
         hr_g1 = exp(estimate_G1),
         lcl_g1 = exp(estimate_G1 - qnorm(.975)*std.error_G1),
         ucl_g1 = exp(estimate_G1 + qnorm(.975)*std.error_G1),
         hr_g0 = exp(estimate_G0),
         lcl_g0 = exp(estimate_G0 - qnorm(.975)*std.error_G0),
         ucl_g0 = exp(estimate_G0 + qnorm(.975)*std.error_G0)) |>
  select(1:5, 19:24, 15, everything()) |>
  arrange(subset, variable, subtype, marker, compartment)

write_csv(select(subs, 1:12), "./Outputs/table_s7.csv")
rm(df, est_mean, fit, imp, imp.long, imp.subset, subgroup.results, study_results, row, sites, models, mi, mv, summaries, uv )
```

## Tables

```{r tables, include=FALSE}
data <- pheno |> 
  mutate(er = ifelse(is.na(er), 9, er),
         age_gp = cut(age_diag, breaks=c(0, 40, 60, 80, Inf), right=FALSE),
         size_gp = cut(size, breaks=c(0, 20, 50, Inf), right=FALSE),
         node_gp = cut(nodes, breaks=c(0, 1, 4, 10, Inf), right=FALSE))
tab1 <- data |> group_by(study) |>
  summarise( cases = n(), age_mean=mean(age_diag), age_min=min(age_diag), age_max=max(age_diag))
write_csv(tab1, "./Outputs/table_1.csv")

myVars <- c("age_gp", "grade", "size_gp", "node_gp", "br_15", "bcss_15", "fu_15")

## Vector of categorical variables that need transformation
catVars <- c("age_gp", "grade", "size_gp", "node_gp", "br_15", "bcss_15")

## Create a TableOne object
tab2 <- CreateTableOne(vars = myVars, data = data, factorVars = catVars, 
                       strata = "er",
                       includeNA = TRUE)

print(tab2, showAllLevels = TRUE, formatOptions = list(big.mark = ","))
rm(tab1, tab2)

```

## Figures

## Figure 1

```{r fig1, echo=FALSE}
#results <- read_csv("./Outputs/results.csv")

limits <- c(0.7, 1.4)
breaks <- c(0.7, 0.8, 0.9, 1, 1.1, 1.2)

plot_fun <- function(var, anal, til, thresh) {
  ggforestplot::forestplot(
    df = filter(results, analysis==anal, variable==var, marker==til, area==thresh),
    name=compartment, 
    estimate = estimate,
    se = std.error,
    colour = adjust,
    logodds= TRUE,
    title = til,
    xlab = "Hazard ratio per standard deviation") +
    scale_color_manual(values=c('cadetblue4','goldenrod4')) +
    scale_x_log10(limits=limits, breaks=breaks) +
    ggforce::facet_col(
      facets = ~subtype,
      scales = "free_y",
      space = "free") +
   theme(legend.position="none")
}

cd8_log_pct <- plot_fun(var="Single marker", anal="Multiple imputation", til="CD8", thresh=0.05) +
  theme(title = element_text(size=10))
cd20_log_pct <- plot_fun(var="Single marker", anal="Multiple imputation", til="CD20", thresh=0.05) +
  theme(title = element_text(size=10))
foxp3_log_pct <- plot_fun(var="Single marker", anal="Multiple imputation", til="FOXP3", thresh=0.05) +
  theme(title = element_text(size=10))
cd163_log_pct <- plot_fun(var="Single marker", anal="Multiple imputation", til="CD163", thresh=0.05) +
  theme(title = element_text(size=10))

fig_1 <- (cd8_log_pct + foxp3_log_pct) / (cd20_log_pct + cd163_log_pct)
ggsave("./Figures/fig_1.jpg", units="cm", height=20, width=17)
rm(cd8_log_pct, cd20_log_pct, cd163_log_pct, foxp3_log_pct)
```

## Figure 2
```{r fig2, echo=FALSE}
limits <- c(0.7, 1.4)
breaks <- c(0.7, 0.8, 0.9, 1, 1.1, 1.2)

mv_plot_function <- function(anal, sub, adj, thresh) {
  ggforestplot::forestplot(
    df = filter(results, variable=="Multi-marker", analysis==anal, 
                subtype==sub, adjust==adj, area==thresh) ,
    name=compartment, 
    colour=marker,
    estimate = estimate,
    se = std.error,
    logodds= TRUE,
    title = paste0(sub, ": ", adj),
    xlab = "Hazard ratio per standard deviation") +
    scale_color_manual(values=c('cadetblue4','goldenrod4', "firebrick4", "darkolivegreen")) +
    scale_x_log10(limits=limits, breaks=breaks) +
    ggforce::facet_col(
      facets = ~marker,
      scales = "free_y",
      space = "free") +
    theme(legend.position="none")
}

mv_mi_par_er0 <- mv_plot_function(anal="Multiple imputation", sub="ER negative", adj="Partial", thresh=0.05) +
  theme(title = element_text(size=10))
mv_mi_par_er1 <- mv_plot_function(anal="Multiple imputation", sub="ER positive", adj="Partial", thresh=0.05) +
  theme(title = element_text(size=10))
mv_mi_full_er0 <- mv_plot_function(anal="Multiple imputation", sub="ER negative", adj="Full", thresh=0.05) +
  theme(title = element_text(size=10))
mv_mi_full_er1 <- mv_plot_function(anal="Multiple imputation", sub="ER positive", adj="Full", thresh=0.05) +
  theme(title = element_text(size=10))

fig_2 <- (mv_mi_par_er0 + mv_mi_full_er0) /  (mv_mi_par_er1 + mv_mi_full_er1)
ggsave("./Figures/fig_2.jpg", units="cm", height=20, width=17)
rm(mv_mi_par_er0, mv_mi_par_er1, mv_mi_full_er0, mv_mi_full_er1)
```

## Figure 3

```{r fig3, include=FALSE}
fig_3 <- ggplot(filter(subs, compartment=="All"), 
                aes(x=hr_g0, y=hr_g1, color=marker, shape=subtype)) +
  geom_point() +
  geom_abline(linetype="dashed") +
  geom_vline(xintercept=1, color='gray80') +
  geom_hline(yintercept=1, color='gray80') +
  scale_x_log10() +
  scale_y_log10() +
  facet_grid(subset~variable) +
  theme_bw() +
  labs(shape="ER status",
       size="P diff < 0.05",
       color="Marker",
       y = "Hazard ratio in first group",
       x = "Hazard ratio in  second group")
fig_3
ggsave("./Figures/fig_3.jpg", units="cm", height=12, width=17)
```

## Supp Figure 1

Boxplot and jitter plot of core tissue area (tumor plus stroma) by study and tumor core diameter

```{r suppfig1, echo=FALSE}
levels <- c('ABC', 'MBC', 'MMH', 'NEA', 'ORI', 'RBC', 'SEA', 'SBC', 'SKK', 'SZB',
            'EPI', 'SGB','BGS', 'CPS', 'EST', 'PSH', 'MAR', 'PBC', 'PLC', 'UNC')
plot_data <- halo |>
  mutate(study = factor(study, levels=levels),
         pct_til_all = area_all_til_halo/(10^4*area_all_mn),
         pct_til_stroma = ifelse(area_stroma_mn==0, NA, area_stroma_til_halo/(area_stroma_mn*10^4)),
         pct_til_tumour = ifelse(area_tumour_mn==0, NA, area_tumour_til_halo/(area_tumour_mn*10^4)),
         pct_artefact = area_artefact_mn/(area_all_mn + area_artefact_mn),
         log_pct_til_all = log(pct_til_all + 0.1),
         log_pct_til_stroma = log(pct_til_stroma + 0.1),
         log_pct_til_tumour = log(pct_til_tumour + 0.1))
fig_s01 <- ggplot(plot_data, aes(x=study, y=area_all_mn, color=as.factor(core_size))) +
  geom_jitter(alpha=0.01) +
  geom_hline(yintercept = c(pi*0.3^2, pi*0.5^2), linetype='dashed') +
  scale_color_manual(values= c('cadetblue3','goldenrod3', "firebrick4")) +
  coord_capped_cart(left="both") +
  geom_boxplot(outliers=FALSE, fill=NA, color='darkblue') + 
  annotate("text", label="0.6 mm cores", x=5, y=1.3) +
  annotate("text", label="1 mm cores", x=15, y=2.1) +
  annotate("text", label="2 mm cores", x=19, y=6.1) +
  theme_classic() +
  theme(legend.position="none",
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x= "Study", y="Core tissue area (sq mm)")
fig_s01
ggsave('./Figures/fig_s01.png', , units="cm", height=10, width=17)
rm(fig_s01)
```

## Sup Figure 2

Boxplot of distribution of TIIC scores (percentage of the tissue area occupied by TIICs) by marker and tumor core size.  Y-axis is log scale.

```{r suppfig2, echo=FALSE}
fig_s02 <- ggplot(plot_data, aes(x=core_size, y=(pct_til_all + 0.1), color=marker)) +
  geom_jitter(alpha=0.05) +
  scale_color_manual(values= c('cadetblue3','goldenrod3', "darkolivegreen", "firebrick4")) +
  scale_y_log10(limits=c(0.1, 100), breaks=c(0.1, 1, 10, 100)) +
  coord_capped_cart(left="both") +
  geom_boxplot(outliers=FALSE, fill=NA, color='darkblue') + 
  facet_wrap(~marker) +
  theme_classic() +
  theme(legend.position="none",
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_blank()) +
  labs(x= "Core size", y="Percent TIL/TAM area")
fig_s02
ggsave("./Figures/fig_s02.png", units="cm", height=10, width=17)
rm(fig_s02)
```
## Sup Figure 3

Scatterplots of pathologist CD8+ TIL scores v automated CD8+ TIL scores (percent tissue area) by tissue compartment

```{r suppfig3, echo=FALSE}

cd8 <- read_csv("./Data/cd8_hra_halo.csv") 
ano <- tibble(compartment=c("All", "Stroma", "Tumour"),
              xtext = c(5, 5, 5),
              ytext = c(1000, 1000, 1000),
              lab = c("Rho=0.76", "Rho=0.62", "Rho=0.71"))
fig_s03 <- ggplot(cd8, aes(x=hra, y=halo)) +
  geom_point(alpha=0.2, color='blue', size=0.5) +
  #geom_smooth( method='lm', se=FALSE, color='grey50', linewidth=0.5) +
  geom_text(data=ano, aes(x=xtext, y=ytext, label=lab), size=3) +
  lemon::coord_capped_cart(left='both', bottom='both') +
  scale_y_log10(breaks=c(0.1, 1, 10, 100, 1000), labels = label_comma(accuracy=1)) +
  scale_x_log10(breaks=c(0.1, 1, 10, 100, 1000), labels = label_comma(accuracy=1)) + 
  facet_wrap(~compartment) + 
  theme_classic() +
  theme(strip.background = element_blank()) +
  labs(x= "Pathologist CD8 cell count per HPF", y = "Automated CD8 cell count")
fig_s03
ggsave("./Figures/fig_s03.png", units="cm", height=10, width=17)
rm(fig_s03, cd8, ano)
```
## Sup Figure 4

Hazard ratio for association between TIIC score and breast cancer specific survival by marker, ER status, tissue compartment and core exclusion threshold. Fully-adjusted single marker models with imputed data.

```{r suppfig4, echo=FALSE}
limits <- c(0.7, 1.4)
breaks <- c(0.7, 0.8, 0.9, 1, 1.1, 1.2)

sens_fun_uv <- function(var, adj, til, anal) {
  ggforestplot::forestplot(
    df = filter(results, variable==var, marker==til, adjust==adj, analysis==anal) ,
    name=compartment, 
    estimate = estimate,
    se = std.error,
    colour = as.factor(area),
    logodds= TRUE,
    title = til,
    xlab = "Hazard ratio per standard deviation") +
    scale_x_log10(limits=limits, breaks=breaks) +
    ggforce::facet_col(
      facets = ~subtype,
      scales = "free_y",
      space = "free") +
    labs(colour = "Tissue area exclusion")
} 

cd8_sens <- sens_fun_uv(var="Single marker", adj="Full", til="CD8", anal="Multiple imputation") +
  theme(title = element_text(size=10))
cd20_sens <- sens_fun_uv(var="Single marker", adj="Full", til="CD20", anal="Multiple imputation")  +
  theme(title = element_text(size=10))
foxp3_sens <- sens_fun_uv(var="Single marker", adj="Full", til="FOXP3", anal="Multiple imputation")  +
  theme(title = element_text(size=10))
cd163_sens <- sens_fun_uv(var="Single marker", adj="Full", til="CD163", anal="Multiple imputation")  +
  theme(title = element_text(size=10))

fig_s04 <- (cd8_sens + foxp3_sens) / (cd20_sens + cd163_sens) +  
  plot_layout(guides = 'collect') 
fig_s04
ggsave("./Figures/fig_s04.png", units="cm", height=20, width=17)
rm(cd8_sens, foxp3_sens, cd20_sens, cd163_sens, fig_s04)
```
## Sup Figure 7

Jitter plot with superimposed boxplot of log(percent artefact+0.1) for each core by marker

```{r suppfig7, echo=FALSE}

fig_s07 <- ggplot(plot_data, aes(x=marker, y=log(pct_artefact*100+0.1))) +
  geom_jitter(alpha=0.02, color='cadetblue4') +
  geom_boxplot(outliers=FALSE, fill=NA, color='darkblue') + 
  coord_capped_cart(left="both") +
  theme_classic() +
  theme(axis.line.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x= "Marker", y = "Log(percent artefact)")
fig_s07
ggsave("./Figures/fig_s07.png", plot=fig_s07, units="cm", height=10, width=17)
rm(fig_s07)

```
## Sup Figure 8

Scatterplot of percent tissue artefact against percent TIIC by marker

```{r suppfig8, echo=FALSE}
fig_s08 <- ggplot(plot_data, aes(x=log(pct_til_all+0.1), y=log(100*pct_artefact+0.1))) +
  geom_point(alpha=0.02, color='cadetblue4') +
  #geom_smooth(method='lm', se=FALSE) +
  lemon::coord_capped_cart(ylim=c(-2.5, 3.0), left='both', bottom='both') +
  facet_wrap(~marker) + theme_classic() +
  theme(strip.background = element_blank()) +
  labs(x= "Log percent TIL", y = "Log percent artefact")
fig_s08
ggsave("./Figures/fig_s08.png", units="cm", height=10, width=17)
rm(fig_s08)
```
## Sup Figure 9

Scatterplot of total tissue area against percent TIIC by marker and tumour core size

```{r suppfig9}

fig_s09 <- ggplot(plot_data, aes(x=log(pct_til_all+0.1), y=area_all_mn)) +
  geom_point(alpha=0.02, color='cadetblue4') +
  geom_smooth(method='lm', se=FALSE, colour='cadetblue4', linewidth=0.5) +
  lemon::coord_capped_cart(ylim=c(-2.5, 3.0), left='both', bottom='both') +
  facet_grid(marker ~ core_size) + theme_classic() +
  theme(strip.background = element_blank()) +
  labs(x= "Log percent TIL", y = "Tissue area mm^2")
fig_s09
ggsave("./Figures/fig_s09.png", units="cm", height=10, width=17)
rm(fig_s09)
```
## supplementary Figure 10 and 11 

Scatterplot of Scaled Schoenfeld residuals against time for grade, number of positive nodes and tumour size in ER-negative Cox regression model

Scatterplot of Scaled Schoenfeld residuals against time for age at diagnosis, grade, number of positive nodes and tumour size in the ER-positive Cox regression model.  age1 and age2 are the transformed variables for age at diagnosis.

```{r suppfig10, echo=FALSE}
imp.long <- read_csv("./Outputs/imputed_data.csv")
df <- imp.long |> filter(er==0, .imp==1) 
fit <-  coxph(Surv(enter, fu_15, br_15) ~ site + age1 + size + nodes + grade,
              data=df)
resid_er0 <- tibble(time = cox.zph(fit)$time, age1 = cox.zph(fit)$y[,2],
                    size = cox.zph(fit)$y[,3], nodes = cox.zph(fit)$y[,4], grade = cox.zph(fit)$y[,5]) |>
  pivot_longer(2:5, values_to="beta", names_to="variable") |>
  filter(beta<600, beta>-600)

fig_s10 <-ggplot(resid_er0, aes(x=time, y = beta)) +
  geom_point(color="blue", alpha=0.2) +
  geom_smooth(method="lm", color="black", se=FALSE) +
  coord_capped_cart(left="both", bottom="both") +
  facet_wrap(~variable, scales="free") +
  theme_classic() +
  theme(strip.background=element_rect(color="white")) +
  labs(x = "Time", y="Beta")
ggsave("./Figures/fig_s10.png", plot=fig_s10, units="cm", height=10, width=17)

df <- imp.long |> filter(er==1, .imp==1)
fit <-  coxph(Surv(enter, fu_15, br_15) ~ site + age1 + age2 + size + nodes + grade,
              data=df)
resid_er1 <- tibble(time = cox.zph(fit)$time, age1 = cox.zph(fit)$y[,2], age2 = cox.zph(fit)$y[,3],
                    size = cox.zph(fit)$y[,4], nodes = cox.zph(fit)$y[,5], grade = cox.zph(fit)$y[,6]) |>
  pivot_longer(2:6, values_to="beta", names_to="variable") |>
  filter(beta<600, beta>-600)

fig_s11 <- ggplot(resid_er1, aes(x=time, y = beta)) +
  geom_point(color="blue", alpha=0.2) +
  geom_smooth(method="lm", color="black", se=FALSE) +
  coord_capped_cart(left="both", bottom="both") +
  facet_wrap(~variable, scales="free") +
  theme_classic() +
  theme(strip.background=element_rect(color="white")) +
  labs(x = "Time", y="Beta")
ggsave("./Figures/fig_s11.png", plot=fig_s11, units="cm", height=10, width=17)
rm(fig_1, fig_2, fig3, fig_s10, fig_s11, imp.long, df, resid_er0, resid_er1)
```
