---
title: "tiic_analysis"
output: html_document
date: "2024-07-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(survival)
library(broom)
library(janitor)
library(meta)
library(mice)
```

## Read in case data and TIIC automated scores

```{r data, include=FALSE}
pheno <- read_csv("./Data/case_data.csv")
halo <- read_csv("./Data/tiic_scores.csv")
```

## Functions for data wrangling and modelling

data_fun generates patient level scores from the automated core level scores as a function of the threshold for excluding cores based on minimum tissue area.  The scores are then merged with the clinical data and the derived age variables created.

```{r functions, include=FALSE}
# Function to generate patient level data for different thresholds of viable core area
data_fun <- function(x, threshold) {
  data <- x |>
    filter(area_all_mn>threshold) |>
    #Generate patient level scores
    group_by(marker, bcac_id) |> 
    summarise(cores=n(),
              area_all = mean(area_all_mn, na.rm=TRUE),
              area_stroma = mean(area_stroma_mn, na.rm=TRUE),
              area_tumour = mean(area_tumour_mn, na.rm=TRUE),
              area_artefact = mean(area_artefact_mn, na.rm=TRUE),
              area_til_all = mean(area_all_til_halo, na.rm=TRUE),
              area_til_stroma = mean(area_stroma_til_halo, na.rm=TRUE),
              area_til_tumour = mean(area_tumour_til_halo, na.rm=TRUE)) |>
    # Generate patient level percent area scores
    mutate(pct_til_all = area_til_all/(10^4*area_all),
           pct_til_stroma = ifelse(area_stroma==0, NA, area_til_stroma/(area_stroma*10^4)),
           pct_til_tumour = ifelse(area_tumour==0, NA, area_til_tumour/(area_tumour*10^4)),
           pct_artefact = area_artefact/(area_artefact + area_all),
           log_pct_til_all = log(pct_til_all + 0.01),
           log_pct_til_stroma = log(pct_til_stroma + 0.01), 
           log_pct_til_tumour = log(pct_til_tumour + 0.01), 
           std_all = scale(log(pct_til_all + 0.1), center=TRUE, scale=TRUE)[,1],
           std_stroma = scale(log(pct_til_stroma + 0.1), center=TRUE, scale=TRUE)[,1],
           std_tumour = scale(log(pct_til_tumour + 0.1), center=TRUE, scale=TRUE)[,1]) |>
    select(bcac_id, marker, cores, 11:13, 18:20) |>
    pivot_wider(id_cols = bcac_id,
                names_from=marker,
                values_from = c(cores, pct_til_all, pct_til_stroma, pct_til_tumour,
                                std_all, std_stroma, std_tumour)) |>
    right_join(filter(pheno, !is.na(er), !is.na(br_15))) |>
    mutate(# Generate PREDICT age variables
      age1 = ifelse(er==1, (age_diag/10)^-2, age_diag),
      age2 = ifelse(er==1, (age_diag/10)^-2*log(age_diag/10), 0)) |>
    select(-er_ast, -er_ari)
  return(data)
}

models_fun <- function(models) {
  summaries <- lapply(models, summary)
  term <- names(df[,6:29])
  estimate <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,1])))
  std.error <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,3])))
  p.value <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,5])))
  n <- as.numeric(unlist(lapply(summaries, function(x) x$n)))
  nevents <- as.numeric(unlist(lapply(summaries, function(x) x$nevent)))
  uv <- tibble(term, estimate, std.error, p.value, n, nevents) |>
    mutate(imputation = "Not applicable",
           variable = "Single marker",
           threshold = threshold,
           subtype = ifelse(er_status==1, "ER positive", "ER negative"),
           analysis = "Complete case analysis")
  return(uv)
}
```

## Single marker complete case analysis to compare percent area with log percent

```{r cca, include=FALSE}
# Vector of thresholds for exluding cores based on minimum tissue area
t_values <- 0.25 #c(0.05, 0.1, 0.15, 0.2, 0.25, 0.3)

# Set up empty tibble for storing results
uni_marker <- tibble(term = character(),
                  estimate = numeric(),
                  std.error = numeric(),
                  p.value = numeric(),
                  subtype = character(),
                  threshold =numeric(),
                  adjust = character(),
                  analysis = character())

for (threshold in t_values) { 
  
  data <- data_fun(halo, threshold=threshold)
  
  for (er_status in 0:1) {
    
    df <- data |> filter(er==er_status) |> mutate(site=ifelse(er==0, site0, site1))
    
    #Partially adjusted single marker models
    models <- lapply(df[,6:29], # [2:25 is the 12 log scores and 12 standardised log scores] 
                     function(x) coxph(Surv(enter, fu_15, br_15) ~ x + age1 + age2 + site, data=df))
    
    uv <- models_fun(models) |> mutate(adjust="Partial")
    uni_marker <-bind_rows(uni_marker, uv)
    
    # Fully adjusted single marker models
    models <- lapply(df[,6:29], 
                     function(x) coxph(Surv(enter, fu_15, br_15) ~ 
                                         x + age1 + age2 + nodes + size + grade + tt(grade) + site, 
                                       tt = function(x, t, ...) { x * log(t)},
                                       data=df))
    uv <- models_fun(models)  |> mutate(adjust="Full")
    uni_marker <- bind_rows(uni_marker, uv) |> 
      mutate(score = ifelse(str_detect(term, "pct"), "Pct", "Std"))
    }
}
rm(data, df)
uni_marker <- uni_marker |>
  mutate(marker= case_when(str_detect(term, "CD8") ~ "CD8",
                           str_detect(term, "CD20") ~ "CD20",
                           str_detect(term, "CD163") ~ "CD163",
                           str_detect(term, "FOXP3") ~ "FOXP3"),
         compartment = case_when(str_detect(term, "all") ~ "All",
                                 str_detect(term, "stroma") ~ "Stroma",
                                 str_detect(term, "tumour") ~ "Tumour"),
         hr = exp(estimate),
         lcl = exp(estimate - qnorm(.975)*std.error),
         ucl = exp(estimate + qnorm(.975)*std.error)) |>
  select(variable, analysis, threshold, subtype, marker, compartment, adjust, score, 
         estimate, std.error, hr, lcl, ucl, p.value, n, nevents) |>
  arrange(variable, analysis, threshold, subtype, marker, compartment, adjust, score)

write_csv(uni_marker, "./Outputs/unimarker_results.csv")

# Count which score (percent area of log percent) is best for each model
uni_marker |> group_by(threshold, subtype, marker, compartment, adjust) |>
  slice(which.min(p.value)) |> ungroup() |> count(score)

uni_marker |> filter(score=="Std") |> group_by(subtype, marker, compartment, adjust, score) |>
  slice(which.min(p.value)) |> ungroup() |> count(threshold)
```

# Analysis with imputation of missing data  ----

Missingness of TIIC scores depends on threshold used for excluding cores with limited tissue area.  Thus data needs to be imputed for each threshold for data exclusion.

After imputation the single marker and multi-marker analyses are run with partial and full adjustment for other porgnostic variables.

```{r impute, include=FALSE}
# Set up empty tibble for storing results
results_i <- tibble(term = character(),
                  estimate = numeric(),
                  std.error = numeric(),
                  p.value = numeric(),
                  subtype = character(),
                  threshold =numeric(),
                  adjust = character(),
                  analysis = character(),
                  score = character())

# Number of data sets to be imputed
iterations <- 10

for (threshold in t_values) { 
  
  mv <- halo |>
    mutate(area_all_mn = ifelse(area_all_mn<threshold, NA, area_all_mn),
           area_stroma_mn = ifelse(area_all_mn<threshold, NA, area_stroma_mn),
           area_tumour_mn = ifelse(area_all_mn<threshold, NA, area_tumour_mn),
           area_all_til_halo = ifelse(area_all_mn<threshold, NA, area_all_til_halo),
           area_stroma_til_halo = ifelse(area_all_mn<threshold, NA, area_stroma_til_halo),
           area_tumour_til_halo = ifelse(area_all_mn<threshold, NA, area_tumour_til_halo)) |>
    #Generate patient level scores
    group_by(marker, bcac_id) |> 
    summarise(cores=n(),
              area_all = mean(area_all_mn, na.rm=TRUE),
              area_stroma = mean(area_stroma_mn, na.rm=TRUE),
              area_tumour = mean(area_tumour_mn, na.rm=TRUE),
              area_artefact = mean(area_artefact_mn, na.rm=TRUE),
              area_til_all = mean(area_all_til_halo, na.rm=TRUE),
              area_til_stroma = mean(area_stroma_til_halo, na.rm=TRUE),
              area_til_tumour = mean(area_tumour_til_halo, na.rm=TRUE)) |>
    mutate(pct_til_all = area_til_all/(10^4*area_all),
           pct_til_stroma = ifelse(area_stroma==0, NA, area_til_stroma/(area_stroma*10^4)),
           pct_til_tumour = ifelse(area_tumour==0, NA, area_til_tumour/(area_tumour*10^4)),
           pct_artefact = area_artefact/(area_artefact + area_all),
           log_pct_til_all = log(pct_til_all + 0.1),
           log_pct_til_stroma = log(pct_til_stroma + 0.1), 
           log_pct_til_tumour = log(pct_til_tumour + 0.1)) |>
    select(bcac_id, marker, cores, 15:17 ) |>
    pivot_wider(id_cols = bcac_id,
                names_from=marker,
                values_from = c(cores, log_pct_til_all, log_pct_til_stroma, log_pct_til_tumour)) |>
    # Generate patient level percent area scores
    right_join(filter(pheno, !is.na(br_15))) |>
    mutate(nodes = log(nodes+0.1),
           size = log(size),
           grade= as.factor(grade),
           er = as.factor(er),
           pr = as.factor(pr),
           her2 = as.factor(her2))
  
  set.seed <- 290962
  imp <- mice(mv, m=iterations, defaultMethod = c("norm", "logreg", "polyreg", "polr"))
  imp.long <- mice::complete(imp,"long",include = T) |>
    mutate(grade = as.numeric(grade),
           size = exp(size),
           nodes = case_when(exp(nodes)-0.1<1e-05 ~ 0, 
                             exp(nodes)-0.1>50 ~ 50,
                             TRUE ~ exp(nodes)-0.1),
           age1 = ifelse(er==1, (age_diag/10)^-2, age_diag),
           age2 = ifelse(er==1, (age_diag/10)^-2*log(age_diag/10), 0),
           std_all_CD8 = scale(log_pct_til_all_CD8, center=TRUE, scale=TRUE)[,1],
           std_stroma_CD8 = scale(log_pct_til_stroma_CD8, center=TRUE, scale=TRUE)[,1],
           std_tumour_CD8 = scale(log_pct_til_tumour_CD8, center=TRUE, scale=TRUE)[,1],
           std_all_CD20 = scale(log_pct_til_all_CD20, center=TRUE, scale=TRUE)[,1],
           std_stroma_CD20 = scale(log_pct_til_stroma_CD20, center=TRUE, scale=TRUE)[,1],
           std_tumour_CD20 = scale(log_pct_til_tumour_CD20, center=TRUE, scale=TRUE)[,1],
           std_all_CD163 = scale(log_pct_til_all_CD163, center=TRUE, scale=TRUE)[,1],
           std_stroma_CD163 = scale(log_pct_til_stroma_CD163, center=TRUE, scale=TRUE)[,1],
           std_tumour_CD163 = scale(log_pct_til_tumour_CD163, center=TRUE, scale=TRUE)[,1],
           std_all_FOXP3 = scale(log_pct_til_all_FOXP3, center=TRUE, scale=TRUE)[,1],
           std_stroma_FOXP3 = scale(log_pct_til_stroma_FOXP3, center=TRUE, scale=TRUE)[,1],
           std_tumour_FOXP3 = scale(log_pct_til_tumour_FOXP3, center=TRUE, scale=TRUE)[,1]) |>
    ungroup() |>
    select(1:3, 20:37, everything())

  for (er_status in 0:1) {
    
    for (iteration in 1:iterations) {
      
      imp.subset <- imp.long.all |> filter(threshold==cutoff, er==er_status, .imp==iteration) |>
        mutate(site=ifelse(er==0, site0, site1))
      
      models <- lapply(imp.subset[,38:49], 
                       function(x) coxph(Surv(enter, fu_15, br_15) ~ x + age1 + age2 + site, 
                                         data=imp.subset))
      summaries <- lapply(models, summary)
      term <- names(imp.subset[,38:49])
      estimate <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,1])))
      std.error <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,3])))
      p.value <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,5])))
      n <- as.numeric(unlist(lapply(summaries, function(x) x$n)))
      nevents <- as.numeric(unlist(lapply(summaries, function(x) x$nevent)))
      uv <- tibble(term, estimate, std.error, p.value, n, nevents) |>
        mutate(n = n,
               nevents = nevents,
               iteration=iteration,
               variable = "Single marker",
               threshold = threshold,
               subtype = ifelse(er_status==1, "ER positive", "ER negative"),
               analysis = "Multiple imputation",
               adjust = "Partial") 
      results_i <-  bind_rows(results_i, uv)
      
      models <- lapply(imp.subset[,38:49], 
                       function(x) coxph(Surv(enter, fu_15, br_15) ~ 
                                           x + age1 + age2 + nodes + size + grade + tt(grade) + site, 
                                         tt = function(x, t, ...) { x * log(t)},
                                         data=imp.subset))
      summaries <- lapply(models, summary)
      term <- names(imp.subset[,38:49])
      estimate <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,1])))
      std.error <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,3])))
      p.value <- as.numeric(unlist(lapply(summaries, function(x) x$coefficients[1,5])))
      n <- as.numeric(unlist(lapply(summaries, function(x) x$n)))
      nevents <- as.numeric(unlist(lapply(summaries, function(x) x$nevent)))
      uv <- tibble(term, estimate, std.error, p.value, n, nevents) |>
        mutate(n = n,
               nevents = nevents,
               iteration=iteration,
               variable = "Single marker",
               threshold = cutoff,
               subtype = ifelse(er_status==1, "ER positive", "ER negative"),
               analysis = "Multiple imputation",
               adjust = "Full") 
      
      results_i <- bind_rows(results_i, uv)
      
      # Partially adjusted multi-marker models
      for (compartment in c("all", "stroma", "tumour")) {
        
        v1 <- paste0("std_", compartment, "_CD8 + ")
        v2 <- paste0("std_", compartment, "_FOXP3 + ")
        v3 <- paste0("std_", compartment, "_CD20 + ")
        v4 <- paste0("std_", compartment, "_CD163 + ")
        
        formula <- paste0("Surv(enter, fu_15, br_15) ~ ", v1, v2, v3, v4, " age1 + age2 +", " site", er_status)
        
        mi <- coxph(as.formula(formula), data=imp.subset)
        mv <- tidy(mi)[1:4, c(1:3,5)] |>
          mutate(iteration=iteration,
                 imputation= "ER included",
                 variable = "Multi-marker",
                 threshold = threshold,
                 subtype = ifelse(er_status==1, "ER positive", "ER negative"),
                 analysis = "Multiple imputation",
                 adjust = "Partial",
                 score = "Std")
        mv$n <- mi$n
        mv$nevents <- mi$nevent
        results_i <- bind_rows(results_i, mv)
        
        # Fully adjusted multi-marker models
        formula <- paste0("Surv(enter, fu_15, br_15) ~ ", v1, v2, v3, v4, 
                          " + age1 + age2 + grade + size + nodes + tt(grade) +", 
                          " site", er_status)
        mi <- coxph(as.formula(formula), 
                    tt = function(x, t, ...) { x * log(t)},
                    data=imp.subset)
        mv <- tidy(mi)[1:4, c(1:3,5)] |>
          mutate(iteration=iteration,
                 imputation= "ER included",
                 variable = "Multi-marker",
                 threshold = threshold,
                 subtype = ifelse(er_status==1, "ER positive", "ER negative"),
                 analysis = "Multiple imputation",
                 adjust = "Full",
                 score = "Std")
        mv$n <- mi$n
        mv$nevents <- mi$nevent
        results_i <- bind_rows(results_i, mv) |>
          mutate(score = ifelse(str_detect(term, "log"), "Log", "Std"))
      }
    }
  }
}
rm(imp, imp.subset, models, mv, mi, uv)

results_i <- results_i |> 
  mutate(marker= case_when(str_detect(term, "CD8") ~ "CD8",
                           str_detect(term, "CD20") ~ "CD20",
                           str_detect(term, "CD163") ~ "CD163",
                           str_detect(term, "FOXP3") ~ "FOXP3"),
         compartment = case_when(str_detect(term, "all") ~ "All",
                                 str_detect(term, "stroma") ~ "Stroma",
                                 str_detect(term, "tumour") ~ "Tumour")) |>
  arrange(subtype, marker, compartment, variable, adjust, score)

```

## Combine the results across analysis of each imputed dataset

Results from analysis of 10 imputed data sets for each exclusion threshold are combined using Rubin's rules and then output saved.

```{r, include=FALSE}
est_mean <- results_i |>
  group_by(score, variable, subtype, marker, compartment, threshold, adjust) |>
  summarise(est_mean=mean(estimate))

imp_results  <-  results_i |>
  mutate(marker= case_when(str_detect(term, "CD8") ~ "CD8",
                           str_detect(term, "CD20") ~ "CD20",
                           str_detect(term, "CD163") ~ "CD163",
                           str_detect(term, "FOXP3") ~ "FOXP3"),
         compartment = case_when(str_detect(term, "all") ~ "All",
                                 str_detect(term, "stroma") ~ "Stroma",
                                 str_detect(term, "tumour") ~ "Tumour")) |>
  left_join(est_mean) |>
  mutate(diff = estimate - est_mean,
         variance=std.error^2) |>
  group_by(score, analysis, imputation, variable, subtype, marker, compartment, threshold, adjust) |>
  summarise(estimate=mean(estimate),
            var_w=mean(variance),
            var_b = mean(diff^2)/9,
            n = round(mean(n)),
            nevents=round(mean(nevents))) |>
  mutate(std.error = (var_w + var_b + var_b/10)^.5,
         statistic = estimate/std.error,
         p.value = 2*pnorm(abs(statistic), lower.tail=F),
         hr = exp(estimate),
         lcl = exp(estimate - qnorm(0.975)*std.error),
         ucl = exp(estimate + qnorm(0.975)*std.error)) |>
  select(-var_w, -var_b, -statistic) |>
  arrange(subtype, marker, compartment, variable, adjust, score)

results <- bind_rows(uni_marker, imp_results)

rm(est_mean, results_i, uni_marker, imp_results)

write_csv(results, "./Outputs/results.csv")

```
## Fraction model variance explained by TIIC scores

The relative variation [21] accounted for by TIIC scores as 
1 – adequacy index = 1 – LRstandard/LRstandard+B
where LR is the likelihood ratio chi-squared statistic, the standard model includes age at diagnosis, ER status, grade, tumor size, number of positive nodes and B denotes the inclusion of TIIC scores as predictors.

Mean value across 10 data sets imputed with the tissue adeauacy threshold set at .25mm2. 

```{r variance, echo=FALSE}

imp.long <- read_csv("./Data/imputed_data.csv") |> filter(threshold==0.25)

adequacy <- tibble(imputation= numeric(),
                   er = numeric(),
                   ll = numeric())

ll_fun <- function(x, y) {
  df <- filter(imp.long, er==x, .imp==y) |>
    mutate(site=ifelse(er==1, site1, site0))
  fit <- coxph(Surv(enter, fu_15, br_15) ~ site + age1 + age2 + size + nodes + grade + tt(grade),
               data=df, 
               tt = function(x, t, ...) { x * log(t)})
  ll_1 <- 2*(fit$loglik[2]-fit$loglik[1])
  fit <- coxph(Surv(enter, fu_15, br_15) ~ site + age1 + age2 + size + nodes + grade + tt(grade)+ 
                 std_all_CD163 + std_all_CD20 +                  std_all_CD8 + std_all_FOXP3,
               data=df, 
               tt = function(x, t, ...) { x * log(t)})
  ll_2 <- 2*(fit$loglik[2]-fit$loglik[1])
  
  ll <- c(ll_1, ll_2)
  return(ll)
}
adequacy <- tibble( imputation = numeric(),
                    er = numeric(),
                    ll_1 = numeric(),
                    ll_2 = numeric())

for (imputation in 1:10) {
  for (er in 0:1) {
    ll <- ll_fun(x=er, y=imputation)
    temp <- tibble(imputation, er, ll_1=ll[1], ll_2=ll[2])
    adequacy <- bind_rows(adequacy, temp)
  }
}

adequacy <- adequacy |>
  mutate(rv = 1 - ll_1/ll_2) |>
  group_by(er) |>
  summarise(rv = mean(rv))
```
## Evaluating heterogeneity between studies

Based on the imputed data when using the 0.25 mm2 threshold
```{r}

# Inter-study heterogeneity----
imp.long <- imp.long |>
  mutate(site=ifelse(er==1, site1, site0))

# Generate empty object to store results
study_results_i <- tibble(study= character(),
                     imputation = numeric(),
                     er = numeric(),
                     term = character(),
                     estimate = numeric(),
                     std.error = numeric())

for (imputation in 1:10) {
  for (er_status in 0:1) {
    studies <- if (er_status==1) {unique(imp.long$site1)} else {unique(imp.long$site0)}
        for (set in studies) {
            df <- imp.long |> filter(er==er_status, .imp==imputation, site==set)
            
            fit <- coxph(Surv(enter, fu_15, br_15) ~ std_all_CD163 + std_all_CD20 + 
                           std_all_CD8 + std_all_FOXP3 + 
                           age1 + age2 + size + nodes + grade + tt(grade),
                         tt = function(x, t, ...) { x * log(t)},
                         data=df)
            
            het <- tidy(fit)[1:4, c(1:3,5)] |>
              mutate(study = set,
                     imputation = imputation,
                     er = er_status)
            study_results_i <- bind_rows(study_results_i, het)
    }
  }
}

est_mean <- study_results_i |>
  mutate(marker= case_when(str_detect(term, "CD8") ~ "CD8",
                           str_detect(term, "CD20") ~ "CD20",
                           str_detect(term, "CD163") ~ "CD163",
                           str_detect(term, "FOXP3") ~ "FOXP3"),
         compartment = case_when(str_detect(term, "all") ~ "All",
                                 str_detect(term, "stroma") ~ "Stroma",
                                 str_detect(term, "tumour") ~ "Tumour")) |>
  group_by(study, er, marker, compartment) |>
  summarise(est_mean=mean(estimate))

study_results  <-  study_results_i |>
  mutate(marker= case_when(str_detect(term, "CD8") ~ "CD8",
                           str_detect(term, "CD20") ~ "CD20",
                           str_detect(term, "CD163") ~ "CD163",
                           str_detect(term, "FOXP3") ~ "FOXP3"),
         compartment = case_when(str_detect(term, "all") ~ "All",
                                 str_detect(term, "stroma") ~ "Stroma",
                                 str_detect(term, "tumour") ~ "Tumour")) |>
  left_join(est_mean) |>
  mutate(diff = estimate - est_mean,
         variance=std.error^2) |>
  group_by(study, er, marker, compartment) |>
  summarise(estimate=mean(estimate),
            var_w=mean(variance),
            var_b = mean(diff^2)/9) |>
  mutate(std.error = (var_w + var_b + var_b/10)^.5,
         statistic = estimate/std.error,
         p.value = 2*pnorm(abs(statistic), lower.tail=F),
         analysis = "Multiple imputation",
         threshold = 0.25,
         adjust = "Full",
         variable = "Multi-marker") |>
  select(analysis, adjust, variable, threshold, marker, compartment, er, study, estimate, std.error, p.value, -var_w, -var_b)

write_csv(study_results, "./Outputs/study_specific_results.csv")

# Meta-analysis of the study specific results
meta_fun <- function(x, y) { 
  df <- study_results |> filter(er==x, marker==y)
  m <- metagen(data=df,
          studlab=study, 
          TE=estimate, seTE=std.error,
          backtransf=TRUE, sm="HR",
          comb.random=FALSE)
  return(c(m$I2, m$pval.Q))
}

het_results <- tibble(marker=character(),
                      er=numeric(),
                      p.het = numeric(),
                      i2 = numeric())

for (marker in c("CD8", "CD20", "CD163", "FOXP3")) {
  for (er in 0:1) {
    test <- meta_fun(x=er, y=marker)
    out <- tibble(marker=marker,
                  er=er,
                  p.het = test[2],
                  i2 = test[1])
    het_results <- bind_rows(het_results, out)
    rm(test, out)
  }
}
  
write_csv(het_results, "./Outputs/het_results.csv")
```

## Figures

```{r, echo=FALSE}
halo <- halo |> 
  mutate(area_all_til_halo = area_stroma_til_halo + area_tumour_til_halo,
         area_all_mn = (area_stroma_mn + area_tumour_mn),
         # Set all areas to missing if no tumour or stroma
         area_all_mn = ifelse(area_all_mn==0, NA, area_all_mn),
         area_artefact_mn = ifelse(area_all_mn==0, NA, area_artefact_mn),
         area_stroma_mn = ifelse(area_all_mn==0, NA, area_stroma_mn),
         area_tumour_mn = ifelse(area_all_mn==0, NA, area_tumour_mn),
         #count_stroma_til_halo= ifelse(area_all_mn==0, NA, count_stroma_til_halo),
         #count_tumour_til_halo= ifelse(area_all_mn==0, NA, count_tumour_til_halo),
         #count_all_til_halo = count_stroma_til_halo + count_tumour_til_halo,
         pct_til_all = area_all_til_halo/(10^4*area_all_mn),
         pct_til_stroma = ifelse(area_stroma_mn==0, NA, area_stroma_til_halo/(area_stroma_mn*10^4)),
         pct_til_tumour = ifelse(area_tumour_mn==0, NA, area_tumour_til_halo/(area_tumour_mn*10^4)),
         log_pct_til_all = log(pct_til_all + 0.1),
         log_pct_til_stroma = log(pct_til_stroma + 0.1),
         log_pct_til_tumour = log(pct_til_tumour + 0.1)) |>
  filter(!is.na(area_stroma_mn)) |>
  left_join(select(pheno, study, bcac_id)) |>
  filter(!is.na(study))
```

## Figure 1

```{r fig1, echo=FALSE}
plot_fun <- function(var, anal, til, thresh) {
  ggforestplot::forestplot(
    df = filter(results, analysis==anal, variable==var, marker==til, threshold==thresh),
    name=compartment, 
    estimate = estimate,
    se = std.error,
    colour = adjust,
    logodds= TRUE,
    title = til,
    xlab = "Hazard ratio per standard deviation") +
    scale_color_manual(values=c('cadetblue4','goldenrod4')) +
    scale_x_log10(limits=limits, breaks=breaks) +
    ggforce::facet_col(
      facets = ~subtype,
      scales = "free_y",
      space = "free") +
   theme(legend.position="none")
}

cd8_log_pct <- plot_fun(var="Single marker", anal="Multiple imputation", til="CD8", thresh=0.25) +
  theme(title = element_text(size=10))
cd20_log_pct <- plot_fun(var="Single marker", anal="Multiple imputation", til="CD20", thresh=0.25) +
  theme(title = element_text(size=10))
foxp3_log_pct <- plot_fun(var="Single marker", anal="Multiple imputation", til="FOXP3", thresh=0.25) +
  theme(title = element_text(size=10))
cd163_log_pct <- plot_fun(var="Single marker", anal="Multiple imputation", til="CD163", thresh=0.25) +
  theme(title = element_text(size=10))

fig1 <- (cd8_log_pct + foxp3_log_pct) / (cd20_log_pct + cd163_log_pct)
fig1
ggsave("./Figures/fig_1.png", units="cm", height=20, width=17)
rm(cd8_log_pct, cd20_log_pct, cd163_log_pct, foxp3_log_pct)
```

## Figure 2
```{r fig2, echo=FALSE}

mv_plot_function <- function(anal, sub, adj, thresh) {
  ggforestplot::forestplot(
    df = filter(results, variable=="Multi-marker", analysis==anal, 
                subtype==sub, adjust==adj, threshold==thresh) ,
    name=compartment, 
    colour=marker,
    estimate = estimate,
    se = std.error,
    logodds= TRUE,
    title = paste0(sub, ": ", adj),
    xlab = "Hazard ratio per standard deviation") +
    scale_color_manual(values=c('cadetblue4','goldenrod4', "firebrick4", "darkolivegreen")) +
    scale_x_log10(limits=limits, breaks=breaks) +
    ggforce::facet_col(
      facets = ~marker,
      scales = "free_y",
      space = "free") +
    theme(legend.position="none")
}

mv_mi_par_er0 <- mv_plot_function(anal="Multiple imputation", sub="ER negative", adj="Partial", thresh=0.25) +
  theme(title = element_text(size=10))
mv_mi_par_er1 <- mv_plot_function(anal="Multiple imputation", sub="ER positive", adj="Partial", thresh=0.25) +
  theme(title = element_text(size=10))
mv_mi_full_er0 <- mv_plot_function(anal="Multiple imputation", sub="ER negative", adj="Full", thresh=0.25) +
  theme(title = element_text(size=10))
mv_mi_full_er1 <- mv_plot_function(anal="Multiple imputation", sub="ER positive", adj="Full", thresh=0.25) +
  theme(title = element_text(size=10))

fig_2 <- (mv_mi_par_er0 + mv_mi_full_er0) /  (mv_mi_par_er1 + mv_mi_full_er1)
fig_2
ggsave("./Figures/fig_2.png", units="cm", height=20, width=17)
rm(mv_mi_par_er0, mv_mi_par_er1, mv_mi_full_er0, mv_mi_full_er1)
```

## Supp Figure 1

Boxplot and jitter plot of core tissue area (tumor plus stroma) by study and tumor core diameter

```{r suppfig1, echo=FALSE}
levels <- c('ABC', 'MBC', 'MMH', 'NEA', 'OFB', 'ORI', 'RBC', 'SEA', 'SBC', 'SKK', 'SZB',
            'EPI', 'SGB','BGS', 'CPS', 'EST', 'PSH', 'GES', 'MAR', 'PBC', 'PLC', 'UNC')
plot_data <- halo |> left_join(pheno) |>
  mutate(study = factor(study, levels=levels))
fig_s01 <- ggplot(plot_data, aes(x=study, y=area_all_mn, color=as.factor(core_size))) +
  geom_jitter(alpha=0.01) +
  geom_hline(yintercept = c(pi*0.3^2, pi*0.5^2), linetype='dashed') +
  scale_color_manual(values= c('cadetblue3','goldenrod3', "firebrick4")) +
  coord_capped_cart(left="both") +
  geom_boxplot(outliers=FALSE, fill=NA, color='darkblue') + 
  annotate("text", label="0.6 mm cores", x=6, y=1.3) +
  annotate("text", label="1 mm cores", x=17, y=2.1) +
  annotate("text", label="2 mm cores", x=21.3, y=6.1) +
  theme_classic() +
  theme(legend.position="none",
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x= "Study", y="Core tissue area (sq mm)")
fig_s1
```

## Sup Figure 2

Boxplot of distribution of TIIC scores (percentage of the tissue area occupied by TIICs) by marker and tumor core size.  Y-axis is log scale.

```{r suppfig2, echo=FALSE}
fig_s02 <- ggplot(plot_data, aes(x=core_size, y=(pct_til_all+0.1), color=marker)) +
  geom_jitter(alpha=0.05) +
  scale_color_manual(values= c('cadetblue3','goldenrod3', "darkolivegreen", "firebrick4")) +
  scale_y_log10(limits=c(0.1, 100), breaks=c(0.1, 1, 10, 100)) +
  coord_capped_cart(left="both") +
  geom_boxplot(outliers=FALSE, fill=NA, color='darkblue') + 
  facet_wrap(~marker) +
  theme_classic() +
  theme(legend.position="none",
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_blank()) +
  labs(x= "Core size", y="Percent TIL/TAM area")
fig_s2
```
## Sup Figure 3

Scatterplots of pathologist CD8+ TIL scores v automated CD8+ TIL scores (percent tissue area) by tissue compartment

```{r suppfig3, echo=FALSE}

cd8 <- read_csv("./Outputs/cd8_hra_halo.csv") 
ano <- tibble(compartment=c("All", "Stroma", "Tumour"),
              xtext = c(5, 5, 5),
              ytext = c(1000, 1000, 1000),
              lab = c("Rho=0.76", "Rho=0.62", "Rho=0.71"))
fig_s3 <- ggplot(cd8, aes(x=hra, y=halo)) +
  geom_point(alpha=0.2, color='blue', size=0.5) +
  #geom_smooth( method='lm', se=FALSE, color='grey50', linewidth=0.5) +
  geom_text(data=ano, aes(x=xtext, y=ytext, label=lab), size=3) +
  lemon::coord_capped_cart(left='both', bottom='both') +
  scale_y_log10(breaks=c(0.1, 1, 10, 100, 1000), labels = label_comma(accuracy=1)) +
  scale_x_log10(breaks=c(0.1, 1, 10, 100, 1000), labels = label_comma(accuracy=1)) + 
  facet_wrap(~compartment) + 
  theme_classic() +
  theme(strip.background = element_blank()) +
  labs(x= "Pathologist CD8 cell count per HPF", y = "Automated CD8 cell count")

fig_s3

```
## Sup Figure 4

Hazard ratio for association between TIIC score and breast cancer specific survival by marker, ER status, tissue compartment and core exclusion threshold. Fully-adjusted single marker models with imputed data.

```{r suppfig4, echo=FALSE}
sens_fun_uv <- function(var, adj, til, anal) {
  ggforestplot::forestplot(
    df = filter(results, variable==var, marker==til, adjust==adj, analysis==anal) ,
    name=compartment, 
    estimate = estimate,
    se = std.error,
    colour = as.factor(threshold),
    logodds= TRUE,
    title = til,
    xlab = "Hazard ratio per standard deviation") +
    scale_x_log10(limits=limits, breaks=breaks) +
    ggforce::facet_col(
      facets = ~subtype,
      scales = "free_y",
      space = "free") +
    labs(colour = "Tissue area exclusion")
} 

cd8_sens <- sens_fun_uv(var="Single marker", adj="Full", til="CD8", anal="Multiple imputation") +
  theme(title = element_text(size=10))
cd20_sens <- sens_fun_uv(var="Single marker", adj="Full", til="CD20", anal="Multiple imputation")  +
  theme(title = element_text(size=10))
foxp3_sens <- sens_fun_uv(var="Single marker", adj="Full", til="FOXP3", anal="Multiple imputation")  +
  theme(title = element_text(size=10))
cd163_sens <- sens_fun_uv(var="Single marker", adj="Full", til="CD163", anal="Multiple imputation")  +
  theme(title = element_text(size=10))

fig_s4 <- (cd8_sens + foxp3_sens) / (cd20_sens + cd163_sens) +  
  plot_layout(guides = 'collect') 
fig_s4
ggsave("./Figures/fig_s04.png", units="cm", height=20, width=17)
rm(cd8_sens, foxp3_sens, cd20_sens, cd163_sens)
```
## Sup Figure 6

Jitter plot with superimposed boxplot of log(percent artefact+0.1) for each core by marker

```{r suppfig6, echo=FALSE}
halo <- halo |>
   mutate(area_all_til_halo = area_stroma_til_halo + area_tumour_til_halo,
         area_all_mn = (area_stroma_mn + area_tumour_mn),
         # Set all areas to missing if no tumour or stroma
         area_all_mn = ifelse(area_all_mn==0, NA, area_all_mn),
         area_artefact_mn = ifelse(area_all_mn==0, NA, area_artefact_mn),
         area_stroma_mn = ifelse(area_all_mn==0, NA, area_stroma_mn),
         area_tumour_mn = ifelse(area_all_mn==0, NA, area_tumour_mn),
         pct_til_all = area_all_til_halo/(10^4*area_all_mn),
         pct_til_stroma = ifelse(area_stroma_mn==0, NA, area_stroma_til_halo/(area_stroma_mn*10^4)),
         pct_til_tumour = ifelse(area_tumour_mn==0, NA, area_tumour_til_halo/(area_tumour_mn*10^4)),
         pct_artefact = area_artefact_mn/(area_all_mn + area_artefact_mn),
         log_pct_til_all = log(pct_til_all + 0.1),
         log_pct_til_stroma = log(pct_til_stroma + 0.1),
         log_pct_til_tumour = log(pct_til_tumour + 0.1)) |>
  filter(!is.na(area_all_mn)) |>
  left_join(select(pheno, study, bcac_id)) |>
  select(study, everything()) |>
  filter(!is.na(study))  

fig_s6 <- ggplot(halo, aes(x=marker, y=log(pct_artefact*100+0.1))) +
  geom_jitter(alpha=0.02, color='cadetblue4') +
  geom_boxplot(outliers=FALSE, fill=NA, color='darkblue') + 
  coord_capped_cart(left="both") +
  theme_classic() +
  theme(axis.line.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x= "Marker", y = "Log(percent artefact)")
fig_s6

```
## Sup Figure 7

Scatterplot of percent tissue artefact against percent TIIC by marker

```{r fig2, echo=FALSE}
fig_s7 <- ggplot(halo, aes(x=log(pct_til_all+0.1), y=log(100*pct_artefact+0.1))) +
  geom_point(alpha=0.02, color='cadetblue4') +
  #geom_smooth(method='lm', se=FALSE) +
  lemon::coord_capped_cart(ylim=c(-2.5, 3.0), left='both', bottom='both') +
  facet_wrap(~marker) + theme_classic() +
  theme(strip.background = element_blank()) +
  labs(x= "Log percent TIL", y = "Log percent artefact")
fig_s7

```
## Sup Figure 8

Scatterplot of total tissue area against percent TIIC by marker and tumour core size

```{r suppfig8}

fig_s8 <- ggplot(halo, aes(x=log(pct_til_all+0.1), y=area_all_mn)) +
  geom_point(alpha=0.02, color='cadetblue4') +
  geom_smooth(method='lm', se=FALSE, colour='cadetblue4', linewidth=0.5) +
  lemon::coord_capped_cart(ylim=c(-2.5, 3.0), left='both', bottom='both') +
  facet_grid(marker ~ core_size) + theme_classic() +
  theme(strip.background = element_blank()) +
  labs(x= "Log percent TIL", y = "Tissue area mm^2")
fig_s8

```
## supplementary Figure 9 and 10 

Scatterplot of Scaled Schoenfeld residuals against time for grade, number of positive nodes and tumour size in ER-negative Cox regression model

Scatterplot of Scaled Schoenfeld residuals against time for age at diagnosis, grade, number of positive nodes and tumour size in the ER-positive Cox regression model.  age1 and age2 are the transformed variables for age at diagnosis.

```{r suppfig9, echo=FALSE}
imp.long <- read_csv("./Outputs/imputed_data_25.csv")
df <- imp.long |> filter(er==0, .imp==1) |> mutate(site=ifelse(er==0, site0, site1))
fit <-  coxph(Surv(enter, fu_15, br_15) ~ site + age1 + size + nodes + grade,
              data=df)
resid_er0 <- tibble(time = cox.zph(fit)$time, age1 = cox.zph(fit)$y[,2],
                    size = cox.zph(fit)$y[,3], nodes = cox.zph(fit)$y[,4], grade = cox.zph(fit)$y[,5]) |>
  pivot_longer(2:5, values_to="beta", names_to="variable") |>
  filter(beta<600, beta>-600)

fig_s9 <-ggplot(resid_er0, aes(x=time, y = beta)) +
  geom_point(color="blue", alpha=0.2) +
  geom_smooth(method="lm", color="black", se=FALSE) +
  coord_capped_cart(left="both", bottom="both") +
  facet_wrap(~variable, scales="free") +
  theme_classic() +
  theme(strip.background=element_rect(color="white")) +
  labs(x = "Time", y="Beta")

df <- imp.long |> filter(er==1, .imp==1) |> mutate(site=ifelse(er==0, site0, site1))
fit <-  coxph(Surv(enter, fu_15, br_15) ~ site + age1 + age2 + size + nodes + grade,
              data=df)
resid_er1 <- tibble(time = cox.zph(fit)$time, age1 = cox.zph(fit)$y[,2], age2 = cox.zph(fit)$y[,3],
                    size = cox.zph(fit)$y[,4], nodes = cox.zph(fit)$y[,5], grade = cox.zph(fit)$y[,6]) |>
  pivot_longer(2:6, values_to="beta", names_to="variable") |>
  filter(beta<600, beta>-600)
fig_s9

fig_s10 <- ggplot(resid_er1, aes(x=time, y = beta)) +
  geom_point(color="blue", alpha=0.2) +
  geom_smooth(method="lm", color="black", se=FALSE) +
  coord_capped_cart(left="both", bottom="both") +
  facet_wrap(~variable, scales="free") +
  theme_classic() +
  theme(strip.background=element_rect(color="white")) +
  labs(x = "Time", y="Beta")
fig_s10
```
